---
title: "Assignment 1"
author: "Natali Tckvitishvili"
date: "`r Sys.Date()`"
output: pdf_document
---
  
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
```


# Task Set 1

For tasks 1.1-1.3, assume you throw 3 dice -- normal dice with 6 sides each ;) 

## Task 1.1

Create a data frame with all possible combinations (outcomes) that can result from throwing all the dice. 
(Each row should represent one possible outcome.) 
Print the first and last 10 rows of the data frame and state how many possible outcomes there are.

Data frame with all possible combinations is created using expand.grid function which creates all the possible combinations of the outcomes.
```{r}
dice_seq <- seq(1:6)
dice_throws <- data.frame(
  expand.grid(dice1 = dice_seq, dice2 = dice_seq, dice3 = dice_seq)
)
```

* top 10 rows

```{r}
head(dice_throws, 10)
```

* last 10 rows

```{r}
tail(dice_throws, 10)
```

* number of possible outcomes

```{r}
nrow(dice_throws)
```

## Task 1.2

Create a data frame showing all possible sums that can result from throwing the three dice along with their probabilities.
Report the results in a summary table (data frame) and a plot (visual graph). 

* summary table

```{r}
dice_throws <- dice_throws %>%
  mutate(possible_sum = dice1 + dice2 + dice3)

total_sum <- nrow(dice_throws) # total number of outcomes to calculate probability

dice_summary <- dice_throws %>%
  group_by(possible_sum) %>%
  summarise(frequency = n()) %>% # count of each sum
  mutate(probability = frequency / total_sum) # probability of each sum

data.frame(dice_summary)
```

* visualization

``` {r}
ggplot(dice_summary, aes(x = as.factor(possible_sum), y = probability)) +
  geom_bar(stat = "identity", fill = "#697188") +
  geom_text(aes(label = round(probability, 3)), vjust = -0.5, size = 3.5) +
  theme_minimal() +
  labs(
    title = "Probability distribution of the outcomes",
    x = "Sum of possible outcomes",
    y = "Probability"
    )
```

## Task 1.3

Compute the probability that the sum is $\geq$ `10`, *given* that at least one of the dice shows a `3`.

```{r}
# assigning variables to avoid hardcode
given_dice <- 3
condition_sum <- 10

given_data <- dice_throws %>%
  filter(dice1 == given_info | dice2 == given_info | dice3 == given_info)
  
condition_data <- given_data %>%
  filter(possible_sum >= condition_sum)

nrow(condition_data) / nrow(given_data)
```

\newpage

# Task Set 2

For Task 2.1-2.3, assume you toss a globe `10` times, leading to either land or water.

## Task 2.1

Compute the probability of all possible numbers of occurrence of land, given the candidate proportion of `.5`.
Report the results in a summary table and a plot and indicate whether the plot shows a probability distribution or a likelihood function.

```{r}
#write code here
```

## Task 2.2

Assume you observe `7` water.
Take the candidate proportions of land `cp =` `r seq(0,1,.1)`.
For each of these candidates, compute the probability of observing `7` water. 
Report the results in a summary table and a plot and indicate whether the plot shows a probability distribution or a likelihood function.

```{r}
#write code here
```

## Task 2.3

For each candidate proportion of land, compute the probability of all possible number of occurrences of land.
Report the results in a summary table, showing the probability distributions as columns and the likelihood functions as rows.

```{r}
#write code here
```


