---
title: "Assignment 5"
author: "Natali Tckvitishvili"
date: "`r Sys.Date()`"
output: pdf_document
execute: 
  cache: true
editor_options: 
  chunk_output_type: console
---

```{r packages, warning=FALSE, message=FALSE}
library(rethinking)
library(dplyr)
library(ggplot2)
library(tidyr)
```

# Task Set 1

Load the data set `RiskyChoice.csv` to solve the Task Set 1. Use the `read_csv2()` function instead of `read_csv()`. 

```{r data}
risk <- read.csv2("/Users/nataly/Management & Technology/Introduction to Bayesian Data Analysis/bayes/bayes/data/RiskyChoice.csv")
head(risk)
summary(risk)
```


## Task 1.1

Create a reduced data table with only one row per subject that shows the number of solved choices problems (`nChoice`) and the number of correct choices (`nCorrect`) for each subject along with the other variables. 
Remove the subjects with missing values. 
Print the data of the first 10 subjects.

```{r}
risk_cut <- risk %>%
  na.omit() %>%
  group_by(Subject) %>%
  summarise(
    nChoice = n(),
    nCorrect = sum(CorrectChoice),
    ageGroup = max(AgeGroup), # they're all the same, so we can take max
    gender = max(Gender),
    numeracy = max(Numeracy))

print(risk_cut, n = 10)
```

## Task 1.2 

Run a Bayesian regression model that predicts `nCorrect` from `Numeracy` using fixed intercepts and fixed slopes.
Standardize the predictor before running the model and compute the WAIC of the model.

```{r}
# make a list
risk_cut_list <- list(
  nCorrect = risk_cut$nCorrect,
  numeracy_s = scale(risk_cut$numeracy),
  subject = as.integer(as.factor(risk_cut$Subject))
)

# plotting to understand the distribution
ggplot(risk_cut, aes(numeracy)) +
  geom_density(color = "#5fc5d5", linewidth = 2)

ggplot(risk_cut, aes(nCorrect)) +
  geom_density(color = "#ca81cb", linewidth = 2)
```

```{r m1, warning=FALSE, message=FALSE, output=FALSE}
#| echo: true
#| eval: true
#| output: false

numeracy_model <- ulam(
  alist(
    nCorrect ~ dnorm(mu, sigma),
    mu <- a + b * numeracy_s,
    a ~ dnorm(0, 0.6),
    b ~ dnorm(0, 0.5),
    sigma ~ dnorm(0, 1)
  ), data = risk_cut_list, chains = 4, cores = 4, log_lik = TRUE
)
```

```{r}
precis(numeracy_model, depth = 2)
WAIC(numeracy_model)
```

## Task 1.3 

Run a Bayesian regression model that predicts `nCorrect` from `Numeracy` using random intercepts and fixed slopes.
Standardize the predictor before running the model and compute the WAIC of the model.

```{r m2, warning=FALSE, message=FALSE, output=FALSE}
#| echo: true
#| eval: true
#| output: false

numeracy_model_subject <- ulam(
  alist(
    nCorrect ~ dnorm(mu, sigma),
    mu <- a[subject] + b * numeracy_s,
    a[subject] ~ dnorm(a_bar, tau_a),
    a_bar ~ dnorm(0, 0.6),
    tau_a ~ dnorm(0, 1),
    b ~ dnorm(1, 0.6),
    sigma ~ dexp(2)
  ), data = risk_cut_list, chains = 4, cores = 4, log_lik = TRUE
)
```

```{r}
precis(numeracy_model_subject, depth = 2)
WAIC(numeracy_model_subject)
compare(numeracy_model, numeracy_model_subject)
```

Looks like if we consider each subject as a group, our model gets better

\newpage

# Task Set 2

## Task 2.1 

Create a data table that entails 10,000 posterior samples (rows) for each subject-specific (columns) intercept. 
Convert the sampled values into probabilities and print the first 10 samples of the first 10 subjects.

```{r}
set.seed(22394)
subject_samples <- extract.samples(numeracy_model_subject, n=10000)$a
probabilities <- as.data.frame(1 / (1 + exp(-subject_samples)))
print(probabilities[1:10, 1:10])
```

## Task 2.2

Use the posterior samples to plot the posterior distribution of all subject-specific intercepts to show the variability in the performance among subjects. 
Use the converted values (probabilities). 

```{r}
probabilities_long <- probabilities %>%
  pivot_longer(cols = everything(), names_to = "subject", values_to = "probability")

ggplot(probabilities_long, aes(x = probability, color = subject)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Supject-specific intercepts",
       x = "Probability",
       y = "Density")
```

Most of the intercepts follow the similar distribution, however, looks like some subjects are different from the others

## Task 2.3 

Consider the following posterior summaries and traceplots. 
Which model was estimated and what might be the cause of the convergence problems?

```{r}
# precis(m3)
# traceplot_ulam(m3, pars = c("mu_a", "tau_a", "mu_b", "tau_b"))
```

The model with random intercepts and slopes was estimated.
Maybe too complicated model was estimated and it would be enough to have either random intercept or random slope model
Other cause could be incorrect specification of the priors for those coefficients
