---
title: "Assignment 4"
author: "Natali Tckvitishvili"
date: "`r Sys.Date()`"
output: pdf_document
execute: 
  cache: true
---

```{r packages, warning=FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(rethinking)
library(scales)
```


```{r data}
heart <- read.csv("heart.csv")
head(heart)
summary(heart)
```


# Task Set 1


## Task 1.1

Run a Bayesian logistic regression model to estimate the risk of men and women to develop a coronary heart disease (TenYearCHD). 
Provide a summary of the posterior distributions.  
What is the average probability of men and women to develop the disease? 

```{r m1, warning=FALSE, message=FALSE, output=FALSE}
#| echo: true
#| eval: true
#| output: false

# gender: 1 is female, 2 is male

heart_list <- list(
  n = nrow(heart),
  gender = heart$male + 1, # re-index as ulam doesn't work with 0
  CHD = heart$TenYearCHD,
  diabetes = heart$diabetes + 1, # for task 2
  age = heart$age,
  avg_age = mean(heart$age)
#  age_women = heart$age[heart$male == 0], # for task 3
#  age_men = heart$age[heart$male == 1]
)

disease_gender <- ulam(
  alist(
    CHD ~ dbinom(n, p), # binomial distribution - success or not
    logit(p) <- a[gender],
    a[gender] ~ dnorm(0, 1)
  ), data = heart_list, chains = 4, cores = 4
)
```

```{r}
traceplot(disease_gender) # looks stable
precis(disease_gender, depth = 2)
inv_logit(coef(disease_gender))
```

Average probabilities to develop the disease for:
* women (gender = 1) ~ 0.003%
* men (gender = 2) ~ 0.005%

# Task 1.2

For the model of `Task 1.1`, visualize the posterior distribution of gender-differences to assess the credibility of the gender difference.

```{r 1.2}
set.seed(1111)
disease_gender_samples <- extract.samples(disease_gender)
posterior <-  tibble(p_female = inv_logit(disease_gender_samples$a[, 1]),
                     p_male = inv_logit(disease_gender_samples$a[, 2]),
                     diff = p_male - p_female)

ggplot(posterior, aes(diff)) + 
  geom_density(color = "#5fc5d5", linewidth = 2) +
   # converting to % scale for easier reading
  scale_x_continuous(labels = percent_format(accuracy = 0.0001)) +
  labs(x = "Disease probability male vs female",
       y = "Density")
```

Most of the probability mass is concentrated between 0.001% and 0.002% difference, therefore, we can assume that there is a high probability that genders differ in chances of developing a coronary heart disease. 
But there might be other confounders, for example, men in general smoke more than women, I suppose smoking increases the risk of heart disease more than just being a man

\newpage

# Task Set 2

## Task 2.1 

Run a Bayesian logistic regression model to estimate the risk of men and women with and without diabetes to develop a coronary heart disease (TenYearCHD). 
Provide a summary of the posterior distributions. 
Does the effect of diabetes differ between men and women?

```{r m2}
#| echo: true
#| eval: true
#| output: false

# diabetes: 1 without, 2 with

disease_diabetes <- ulam(
  alist(
    CHD ~ dbinom(n, p), # binomial distribution - success or not
    logit(p) <- a[gender, diabetes],
    matrix[gender, diabetes]:a ~ dnorm(0, 1)
  ), data = heart_list, chains = 4, cores = 4
)
```

```{r 2.1}
traceplot(disease_diabetes) # looks stable
precis(disease_diabetes, depth = 3)

# renaming columns to make analysis easier
coef_diabetes <- inv_logit(coef(disease_diabetes))
names(coef_diabetes) <- c("female_without", "male_without", "female_with", "male_with")
coef_diabetes
```

Among people who have diabetes the difference is not as big as among those who do not have it:
* women with diabetes: ~ 0.011%
* men with diabetes: ~ 0.014%

* women without diabetes: ~ 0.0029%
* men without diabetes: ~ 0.0044%

# Task 2.2

For the model of `Task 2.1`, visualize the posterior distributions of each group in one plot to better assess the credibility of the group differences. 

```{r 2.2}
set.seed(1111)
disease_diabetes_samples <- data.frame(extract.samples(disease_diabetes))
names(disease_diabetes_samples) <- c("female_without", "male_without", "female_with", "male_with")

posterior_diabetes <-  data.frame(
  female_without = inv_logit(disease_diabetes_samples$female_without),
  male_without = inv_logit(disease_diabetes_samples$male_without),
  female_with = inv_logit(disease_diabetes_samples$female_with),
  male_with = inv_logit(disease_diabetes_samples$male_with)
  ) %>%
  pivot_longer(cols = female_without:male_with, names_to = "Type", values_to = "Probability")

ggplot(posterior_diabetes, aes(Probability, color = Type)) +
  geom_density(alpha = .3, linewidth = 2) +
  scale_x_continuous(labels = percent_format(accuracy = 0.0001)) +
  labs(x = "Probability",
       y = "Density",
       title = "Probability of developing CHD with and without diabetes"
       )
```

According to the graph, men are under higher risk than women to develop coronary heart disease both with and without diabetes
Besides, both genders with diabetes have significantly higher chances to develop CHD than those without.

\newpage

# Task Set 3

## Task 3.1 

Run a Bayesian logistic regression model to estimate the effect of age on the risk of developing a coronary heart disease (TenYearCHD), separately for women and men. 
Ensure that the regression intercept represents the risk of women and men with average age.
Provide a summary of the posterior distributions. 

```{r m, warning=FALSE, message=FALSE}
#| echo: true
#| eval: true
#| output: false
women_list <- list(
  n = nrow(heart[heart$male == 0, ]),
  CHD = heart$TenYearCHD[heart$male == 0],
  age_women = heart$age[heart$male == 0],
  avg_age_women = mean(heart$age[heart$male == 0])
)

disease_age_women <- ulam(
  alist(
    CHD ~ dbinom(n, p),
    logit(p) <- a_1 + b_1 * (age_women - avg_age_women), # standardizing around the mean
    a_1 ~ dnorm(10, 5),
    b_1 ~ dnorm(1, 0.1)
  ), data = women_list, chains = 4, cores = 4
)
```

```{r 3.1}
traceplot(disease_age_women)
precis(disease_age_women, depth = 2)
inv_logit(coef(disease_age_women))
```

```{r m4, warning=FALSE, message=FALSE}
#| echo: true
#| eval: true
#| output: false
men_list <- list(
  n = nrow(heart[heart$male == 1, ]),
  CHD = heart$TenYearCHD[heart$male == 1],
  age_men = heart$age[heart$male == 1],
  avg_age_men = mean(heart$age[heart$male == 1])
)

disease_age_men <- ulam(
  alist(
    CHD ~ dbinom(n, p),
    logit(p) <- a_1 + b_1 * (age_men - avg_age_men), # standardizing around the mean
    a_1 ~ dnorm(10, 5),
    b_1 ~ dnorm(1, 0.1)
  ), data = men_list, chains = 4, cores = 4
)
```

```{r m4.1}
traceplot(disease_age_men)
precis(disease_age_men, depth = 2)
inv_logit(coef(disease_age_men))
```
## Task 3.2 

For the model of `Task 3.1`, visualize the posterior distribution of differences in the age effect between women and men. 
Does age increase the risk of developing the disease and does this effect differ between women and men? 

```{r 3.2}
set.seed(1111)
women_age_samples <- data.frame(extract.samples(disease_age_women))
head(women_age_samples)

men_age_samples <- data.frame(extract.samples(disease_age_men))
head(men_age_samples)

posterior_age <- data.frame(
  female = women_age_samples$b_1,
  male = men_age_samples$b_1
  ) %>%
  pivot_longer(cols = female:male, names_to = "Gender", values_to = "Coef")

ggplot(posterior_age, aes(Coef, color = Gender)) +
  geom_density(alpha = .3, linewidth = 2) +
  scale_x_continuous(labels = percent_format(accuracy = 0.01)) +
  labs(x = "Probability",
       y = "Density",
       title = "Increase in chances of getting CHD with increasing age"
       )
```

With increasing age, women tend to have higher risk of developing CHD than men
That, honestly, doesn't sound right, but I don't know how to change the model :(